### 2015-02-27 examples/qval.nonp.rmd
split ash into ash and ash.workhorse

In our first simulation, half the $\beta_j$ are null, and the other half are $\beta_j \sim N(0,2^2)$. Then $\hat\beta_j \sim N(\beta_j, 1)$. We apply `qvalue`,`ash`,`locfdr` and `mixFdr` to estimate the $q$-values. (The last three methods all estimate local fdr values; we convert these to corresponding estimated $q$ values using the `ashr` function `qval.from.lfdr`, which, for each observation $j$, computes the $q$ value as the average of the lfdr values more significant than observation $j$.)
```{r results='hide'}
  set.seed(100)
  test = simdata(10000, 5000, 0, 2, 1) 
  test.ash = ash(test$betahat,test$betahatsd,method="fdr")
  z = test$betahat/test$betahatsd
  p = pchisq(z^2, df=1,lower.tail=FALSE)
  test.qv = qvalue(p)
  test.locfdr = locfdr(z,nulltype=0,plot=0)
  test.locfdr.qval = qval.from.lfdr(test.locfdr$fdr)
  test.mixfdr = mixFdr(z,theonull=TRUE,noiseSD=1,plots=FALSE)
  test.mixfdr.qval = qval.from.lfdr(test.mixfdr$fdr)
```

@@

Now we perform a "sanity check" for each method, comparing its $q$-values against the
nonparametric $q$ values obtained using the same estimate of $\pi_0$. Both `ash` and `qvalue`
clearly pass this sanity check (for `qvalue`, this is almost tautological, since it effectively uses the non-parametric procedure). For `locfdr`, it also passes the sanity check except for
some artifacts near $p=1$ that we do not fully understand. For `mixfdr` there seems to be a systematic deviation from the non-parametric method that suggests...
```{r}
  par(mfcol=c(2,2))
  plot(qval.nonp(p,get_pi0(test.ash)),test.ash$qvalue, xlab="nonparametric q-values (with pi0 estimated by ash)", ylab = "ash q-values",main="ash")
  abline(a=0,b=1,col=2,lwd=2)

  plot(qval.nonp(p,test.qv$pi0),test.qv$qvalues, xlab="nonparametric q-values (with pi0 estimated by qvalue)", ylab = "qvalue q-values",main="qvalue")
  abline(a=0,b=1,col=2,lwd=2)

 plot(qval.nonp(p,test.mixfdr$pi[1]),test.mixfdr.qval, xlab="nonparametric q-values (with pi0 estimated by mixfdr)", ylab = "mixfdr q-values",main="mixfdr")
  abline(a=0,b=1,col=2,lwd=2)

 plot(qval.nonp(p,test.locfdr$fp0[1,3]),test.locfdr.qval, xlab="nonparametric q-values (with pi0 estimated by locfdr)", ylab = "locfdr q-values", main="locfdr")
  abline(a=0,b=1,col=2,lwd=2) 
  par(mfcol=c(1,1))
```

In this second simulation, half the $\beta_j$ are null, and the alternatives
are $\beta_j \sim N(2,1)$. Note that this deviates quite a bit from the `ash` assumption that
the non-zero $\beta_j$ will have a unimodal distribution centered on 0. This presumably
accounts for the slightly worse correspondance between the `ash` and non-parametric $q$-value compared with the previous simulation.
```{r}
  set.seed(100)
  test = simdata(10000, 5000, 2, 1, 1) 
  test.ash = ash(test$betahat,test$betahatsd,method="fdr")
  z = test$betahat/test$betahatsd
  p = pchisq(z^2, df=1,lower.tail=FALSE)
  plot(qval.nonp(p,get_pi0(test.ash)),test.ash$qvalue, xlab="nonparametric q-values (ash estimate of pi0)", ylab = "ash q-values")
  abline(a=0,b=1,col=2,lwd=2)
```

>>>>

In our first simulation, half the $\beta_j$ are null, and the other half are $\beta_j \sim N(0,2^2)$. Then $\hat\beta_j \sim N(\beta_j, 1)$. We apply `qvalue`,`ash`,`locfdr` and `mixFdr` to estimate the $q$-values. (The last three methods all estimate local fdr values; we convert these to corresponding estimated $q$ values using the `ashr` function `qval.from.lfdr`, which, for each observation $j$, computes the $q$ value as the average of the lfdr values more significant than observation $j$.)
```{r results='hide'}
  set.seed(100)
  test = simdata(10000, 5000, 0, 2, 1) 
  test.ash = ash(test$betahat,test$betahatsd)
  test.ash.hu = ash(test$betahat,test$betahatsd,mixcompdist="halfuniform")
  z = test$betahat/test$betahatsd
  p = pchisq(z^2, df=1,lower.tail=FALSE)
  test.qv = qvalue(p)
  test.locfdr = locfdr(z,nulltype=0,plot=0)
  test.locfdr.qval = qval.from.lfdr(test.locfdr$fdr)
  test.mixfdr = mixFdr(z,theonull=TRUE,noiseSD=1,plots=FALSE)
  test.mixfdr.qval = qval.from.lfdr(test.mixfdr$fdr)
```

@@

Now we perform a "sanity check" for each method, comparing its $q$-values against the
nonparametric $q$ values obtained using the same estimate of $\pi_0$. Both `ash` and `qvalue`
clearly pass this sanity check (for `qvalue`, this is almost tautological, since it effectively uses the non-parametric procedure). For `locfdr`, it also passes the sanity check except for
some artifacts near $p=1$ that we do not fully understand. For `mixfdr` there seems to be a systematic deviation from the non-parametric method that suggests...
```{r}

  res = rbind(data.frame(method="qvalue",q = test.qv$qvalues, qnonp = qval.nonp(p,test.qv$pi0)),
         data.frame(method="mixfdr",q = test.mixfdr.qval, qnonp = qval.nonp(p,test.mixfdr$pi[1])),
         data.frame(method="locfdr",q = test.locfdr.qval, qnonp = qval.nonp(p,test.locfdr$fp0[1,3])),

         data.frame(method="ashr",q = test.ash$qvalue, qnonp = qval.nonp(p,get_pi0(test.ash)))
         )
  library(ggplot2)
  ggplot(res, aes(x=q, y=qnonp)) + facet_grid(.~method) + geom_point(shape=1) +xlim(0,1) +ylim(0,1) + xlab("qvalue computed by method") +ylab("qvalue computed \n non-parametrically") +geom_abline(slope=1,intercept=0,color=2) + coord_equal() + theme(axis.text.x  = element_text(size=8), axis.text.y  = element_text(size=8))
  ggsave("qval_comparison.png", width=8, height=2, dpi=100)

```

In this second simulation, half the $\beta_j$ are null, and the alternatives
are $\beta_j \sim N(+-2,1)$. Note that this deviates quite a bit from the `ash` assumption that
the non-zero $\beta_j$ will have a unimodal distribution centered on 0. This presumably
accounts for the slightly worse correspondance between the `ash` and non-parametric $q$-value compared with the previous simulation.
```{r}
  set.seed(100)
  test = rbind(data.frame(simdata(5000, 2500, 2, 1, 1)), data.frame(simdata(5000,2500,-2,1,1))) 
  test.ash = ash(test$betahat,test$betahatsd)

  z = test$betahat/test$betahatsd
  p = pchisq(z^2, df=1,lower.tail=FALSE)
  test.qv = qvalue(p)
  test.locfdr = locfdr(z,nulltype=0,plot=0)
  test.locfdr.qval = qval.from.lfdr(test.locfdr$fdr)
  test.mixfdr = mixFdr(z,theonull=TRUE,noiseSD=1,plots=FALSE)
  test.mixfdr.qval = qval.from.lfdr(test.mixfdr$fdr)
  kable(data.frame(method=c("qvalue","ash","locfdr","mixfdr"),pi0=round(c(test.qv$pi0,get_pi0(test.ash),test.locfdr$fp0[1,3],test.mixfdr$pi[1]),digits=2)))

  res = rbind(data.frame(method="qvalue",q = test.qv$qvalues, qnonp = qval.nonp(p,test.qv$pi0)),
         data.frame(method="mixfdr",q = test.mixfdr.qval, qnonp = qval.nonp(p,test.mixfdr$pi[1])),
         data.frame(method="locfdr",q = test.locfdr.qval, qnonp = qval.nonp(p,test.locfdr$fp0[1,3])),

         data.frame(method="ashr",q = test.ash$qvalue, qnonp = qval.nonp(p,get_pi0(test.ash)))
         )
  library(ggplot2)
  ggplot(res, aes(x=q, y=qnonp)) + facet_grid(.~method) + geom_point(shape=1) +xlim(0,1) +ylim(0,1) + xlab("qvalue computed by method") +ylab("qvalue computed \n non-parametrically") +geom_abline(slope=1,intercept=0,color=2) + coord_equal() + theme(axis.text.x  = element_text(size=8), axis.text.y  = element_text(size=8))
  ggsave("qval_comparison_nonUA.png", width=8, height=2, dpi=100)

```
